---
title: "Part 2 Homework"
output: html_document
---
```{r version-and-curren-date}
R.version.string
Sys.Date()
```

```{r read-gardasil}
fn <- "gardasil.csv"
ga <- read.csv(fn)
head(ga)
tail(ga)
```
1. Create factors for AgeGroup, Race, Completed, Location.

```{r-factors}
ga$AgeGroup.factor <- factor(ga$AgeGroup,levels = 0:1, labels = c("11-17","18-26"))
ga$Race.factor <- factor(ga$Race,levels = 0:3, labels = c("white","black","Hispanic","other/unknown"))
ga$Completed.factor <- factor(ga$Completed, levels = 0:1, labels = c("no","yes"))
ga$Location.factor <- factor(ga$Location,levels = 1:4, labels = c("Odenton","White Marsh","Johns Hopkins Outpatient Center","Bayview"))
```


2. Report which variables have missing data and how many of these values are missing.

```{r-missing-values}
addmargins(table(ga$Age,useNA="always"))
addmargins(table(ga$AgeGroup,useNA="always"))
addmargins(table(ga$Race,useNA="always"))
addmargins(table(ga$Shots,useNA="always"))
addmargins(table(ga$Completed,useNA="always"))
addmargins(table(ga$InsuranceType,useNA="always"))
addmargins(table(ga$MedAssist,useNA="always"))
addmargins(table(ga$Location,useNA="always"))
addmargins(table(ga$LocationType,useNA="always"))
addmargins(table(ga$PracticeType,useNA="always"))
```

3. What proportion of patients completed all three shots?

```{r-proportion}
tb <- table(ga$Completed.factor)
prop.table(tb)
pct <- round(100*prop.table(tb))
pct.sign <- "%"
paste(names(pct), " ", pct,pct.sign,sep = "")
```
33% of patients completed all shots

4. Draw a bar chart showing the percentage of patients at each of the four locations.

```{r-barplot, fig.width= 10}
tbloc <- table(ga$Location.factor)
locpct <- round(100*prop.table(tbloc))
barplot(locpct,xlab = "Location", ylab = "Percent of Patients")
```

5. Use a crosstabulation to compare Age to AgeGroup, and Location to LocationType. Are the recodings into AgeGroup and LocationType done properly?

```{r-crosstabulation}
table(ga$Age,ga$AgeGroup.factor)

ga$Locationtype.factor <- factor(ga$LocationType,levels = 0:1, labels = c("suburban","urban"))
table(ga$Location.factor,ga$Locationtype.factor)
```

6. Create a new variable that combines the race categories into white, and non-white. Calculate the proportion of white patients at each of the four locations.

```{r-new-variable}
ga$nonwhite <- as.numeric(ga$Race > 0)
ga$nonwhite.factor <- factor(ga$nonwhite,levels = 0:1, labels = c("white","nonwhite"))

prop <-table(ga$Location.factor,ga$nonwhite.factor)
addmargins(prop)
round(100*prop.table(prop,margin = 1),2)
```


7. Does the likelihood of completing all four shots vary by AgeGroup, Race, or Location? Calculate the appropriate percentages and relative risks.

```{r-oddsratio}
library(epitools)
oddsratio(ga$AgeGroup.factor,ga$Completed.factor)$measure
oddsratio(ga$Race.factor,ga$Completed.factor)$measure
oddsratio(ga$Location.factor,ga$Completed.factor)$measure
```

```{r-riskratio}
riskratio(ga$AgeGroup.factor,ga$Completed.factor)$measure
riskratio(ga$Race.factor,ga$Completed.factor)$measure
riskratio(ga$Location.factor,ga$Completed.factor)$measure
```

The likelihood of completing all shots is variable accross all listed categories.


8. Place all the key results into a Word document, PowerPoint presentation, as a PDF file or as an HTML file. Turn it in by email. Use the subject line: "Introduction to R, (your name), Part 2 Homework"
