---
title: "Part 2 Homework"
output: html_document
---
```{r version-and-curren-date}
R.version.string
Sys.Date()
```

```{r read-gardasil}
#Reading in data from .csv file
fn <- "gardasil.csv"
ga <- read.csv(fn)
#Head of table
head(ga)
#End of table
tail(ga)
```
1. Create factors for AgeGroup, Race, Completed, Location.

```{r-factors}
#AgeGroup Factor
ga$AgeGroup.factor <- factor(ga$AgeGroup,levels = 0:1, labels = c("11-17","18-26"))
#Race Factor
ga$Race.factor <- factor(ga$Race,levels = 0:3, labels = c("white","black","Hispanic","other/unknown"))
#Completed Factor
ga$Completed.factor <- factor(ga$Completed, levels = 0:1, labels = c("no","yes"))
#Location Factor
ga$Location.factor <- factor(ga$Location,levels = 1:4, labels = c("Odenton","White Marsh","Johns Hopkins Outpatient Center","Bayview"))
```
Variable factors


2. Report which variables have missing data and how many of these values are missing.

```{r-missing-values}
#No missing data for Age
addmargins(table(ga$Age,useNA="always"))
#No missing data for AgeGroup
addmargins(table(ga$AgeGroup,useNA="always"))
#No missing data for Race
addmargins(table(ga$Race,useNA="always"))
#No missing data for Shots
addmargins(table(ga$Shots,useNA="always"))
#No missing data for Completed
addmargins(table(ga$Completed,useNA="always"))
#No missing data for InsuranceType
addmargins(table(ga$InsuranceType,useNA="always"))
#No missing data for MedAssist
addmargins(table(ga$MedAssist,useNA="always"))
#No missing data for Location
addmargins(table(ga$Location,useNA="always"))
#No missing data for LocationType
addmargins(table(ga$LocationType,useNA="always"))
#No missing data for PracticeType
addmargins(table(ga$PracticeType,useNA="always"))
```
No variables had missing data

3. What proportion of patients completed all three shots?

```{r-proportion}
#Proportion of patient's which completed shots
tb <- table(ga$Completed.factor)
prop.table(tb)

#Converting proportion to percent
pct <- round(100*prop.table(tb))
pct.sign <- "%"
paste(names(pct), " ", pct,pct.sign,sep = "")
```
33% of patients completed all shots

4. Draw a bar chart showing the percentage of patients at each of the four locations.

```{r-barplot, fig.width= 10}
#Bar plot with percentage of patients at the locations (fig.width = 10)
tbloc <- table(ga$Location.factor)
locpct <- round(100*prop.table(tbloc))
barplot(locpct,xlab = "Location", ylab = "Percent of Patients")
```

Barplot with percentage of patient's at each location

5. Use a crosstabulation to compare Age to AgeGroup, and Location to LocationType. Are the recodings into AgeGroup and LocationType done properly?

```{r-crosstabulation}
#Crosstabulation for Age to AgeGroup
table(ga$Age,ga$AgeGroup.factor)

#Crosstabulation for Location to LocationType
ga$Locationtype.factor <- factor(ga$LocationType,levels = 0:1, labels = c("suburban","urban"))
table(ga$Location.factor,ga$Locationtype.factor)
```

The recodings do appear to be done properly

6. Create a new variable that combines the race categories into white, and non-white. Calculate the proportion of white patients at each of the four locations.

```{r-new-variable}
#Creating new variable
ga$nonwhite <- as.numeric(ga$Race > 0)
#Variable factors
ga$nonwhite.factor <- factor(ga$nonwhite,levels = 0:1, labels = c("white","nonwhite"))
#Calculation proportion at each location
prop <-table(ga$Location.factor,ga$nonwhite.factor)
addmargins(prop)
round(100*prop.table(prop,margin = 1),2)
```



7. Does the likelihood of completing all four shots vary by AgeGroup, Race, or Location? Calculate the appropriate percentages and relative risks.

```{r-oddsratio}
#Loading epitools library
library(epitools)
#Odds of completing shots by age group
oddsratio(ga$AgeGroup.factor,ga$Completed.factor)$measure
#Odds of completing shots by race
oddsratio(ga$Race.factor,ga$Completed.factor)$measure
#Odds of completing shots by location
oddsratio(ga$Location.factor,ga$Completed.factor)$measure
```

```{r-riskratio}
#Risk of completing shots by age group
riskratio(ga$AgeGroup.factor,ga$Completed.factor)$measure
#Risk of completing shots by race
riskratio(ga$Race.factor,ga$Completed.factor)$measure
#Risk of completing shots by location
riskratio(ga$Location.factor,ga$Completed.factor)$measure
```

The likelihood of completing all shots is variable accross all listed categories.


8. Place all the key results into a Word document, PowerPoint presentation, as a PDF file or as an HTML file. Turn it in by email. Use the subject line: "Introduction to R, (your name), Part 2 Homework"
