---
title: "Homework Part 3"
output: html_document
---

```{r r-version-and-current-date}
R.version.string
Sys.Date()
```

```{r housing-data}
#Reading in data
house <- read.table(file="housing.txt",header=TRUE,as.is=TRUE)
head(house)
tail(house)
```

1. Convert the asterisks in the TAX variable to missing. How many missing values are there for AGE and for TAX?

```{r asterisks-to-NA}
#Converting asterisks to NAs
house$age.years <- as.numeric(house$AGE)
summary(house$age.years)

house$tax.value <- as.numeric(house$TAX)
summary(house$tax.value)
```

There are 49 missing values for age, and 10 missing values for taxes

2. Create factors for NE, CUST, and COR. Draw bar charts for each of these factors.

```{r factors}
#NE factor
house$NE.factor <- factor(house$NE, levels = 0:1, labels = c("No","Yes"))

#CUST factor
house$CUST.factor <- factor(house$CUST, levels = 0:1, labels = c("No","Yes"))

#COR factor
house$COR.factor <- factor(house$COR, levels = 0:1, labels = c("No","Yes"))
```

```{r bar-charts}
#NE barplot
tabNE <- table(house$NE.factor)
barplot(tabNE, xlab ="In the Northeast", ylab = "Frequency")

#CUST barplot
tabCUST <- table(house$CUST.factor)
barplot(tabCUST, xlab ="Custom", ylab = "Frequency")

#COR barplot
tablCOR <- table(house$COR.factor)
barplot(tablCOR, xlab="Corner",ylab="Frequency")
```

3. Find the largest house (biggest SQFT) in the data set. Is the largest house also the most expensive house?

```{r max-sqft}
#Summary of SQFT
summary(house$SQFT)
#Largest house
house[which.max(house$SQFT), ]
```
```{r}
#Summary of cost
summary(house$PRICE)
#Most expensive house
house[which.max(house$PRICE), ]
```
The largest house is not the most expensive house

4. Calculate frequency counts for FEATS. Are there any houses with the no features? Are there any houses with every possible feature?

```{r frequencies}
#FEATS frequencies
table(house$FEATS)
hist(house$FEATS, xlab = "Number of Features", ylab = "Frequency")
```

Two houses have no features
No house has every feature (Max is 8, Total possible is 11)

5. Evaluate the relationship between PRICE and SQFT using a scatterplot. Include a smooth curve. Do larger houses tend to cost more?

```{r scatterplot}
#Price vs Square Feet
plot(house$PRICE,house$SQFT, xlab = "Price", ylab = "Square Feet")
smooth.curve <- lowess(house$PRICE,house$SQFT)
lines(smooth.curve)
```

Yes, larger houses tend to cost more

6. Evaluate the relationship between CUST and PRICE using a boxplot. Calculate the mean price and standard deviation by CUST.

```{r boxplot}
#Boxplot CUST vs PRICE
boxplot(PRICE~CUST.factor, data=house)

#Mean
by(house$PRICE,house$CUST.factor,mean,na.rm=TRUE)

#SD
by(house$PRICE,house$CUST.factor,sd,na.rm=TRUE)


#Mean and SD
grp.means <- by(house$PRICE,house$CUST.factor,mean,na.rm=TRUE)
grp.stdev <- by(house$PRICE,house$CUST.factor,sd,na.rm=TRUE)
colon <- ": "
plus.minus <- "+/-"
paste(names(grp.means),colon,round(grp.means,1),
      plus.minus,round(grp.stdev,1),sep="")



```


7. What is the difference in average prices between northeast houses and other houses? What is the difference in average taxes?

```{r averages-by-NE}
#Mean price by location
by(house$PRICE,house$NE.factor,mean,na.rm=TRUE)
price.means <- by(house$PRICE,house$NE.factor,mean,na.rm=TRUE)
#Mean tax by location
by(house$tax.value,house$NE.factor,mean,na.rm=TRUE)
tax.means <- by(house$tax.value,house$NE.factor,mean,na.rm=TRUE)
#Average difference taxes
tax.dif <- tax.means[["Yes"]] - tax.means[["No"]]
paste("Difference: ",round(tax.means[["Yes"]],2),"-",round(tax.means[["No"]],2),"=",round(tax.dif,2))
#Average differencec price
price.dif <- price.means[["Yes"]] - price.means[["No"]]
paste("Difference: ",round(price.means[["Yes"]],2),"-",round(price.means[["No"]],2),"=",round(price.dif,2))
```

8. Are custombuilt houses more likely to appear on corner lots? Calculate the percentages and compute a relative risk.
```{r}
house$COR.factor.2 <- factor(house$COR, levels = 0:1, labels = c("Not-Corner","Corner"))
house$CUST.factor.2 <- factor(house$CUST, levels = 0:1, labels = c("Not-Custom","Custom"))

library(epitools)
#Percentages
tb <- table(house$COR.factor.2,house$CUST.factor.2)
round(100*prop.table(tb,margin=2),2)

#Relative risk
riskratio(tb)$measure
```

Custom houses are not more likely to appear on corner lots


9. Place all the key results into a Word document, add a bit of commentary, and turn it in.