---
title: "Part 1 Homework"
output: html_document
---

```{r}
R.version.string
Sys.Date()
```


```{r reading in data}
sleepdata <- "http://www.statsci.org/data/general/sleep.txt"
sleeptable <- read.table(file=sleepdata,header=TRUE)
head(sleeptable)
tail(sleeptable)
```
1. Notice that there is a huge range in body weight. Display the information for the smallest and the largest mammals.

```{r finding smallest and largest BodyWt}
summary(sleeptable$BodyWt)
```

Min = .005
Max = 6654.00

```{r min}
sleeptable[which(sleeptable$BodyWt==.005), ]
```

```{r max}
sleeptable[which(sleeptable$BodyWt==6654.00), ]
```

2. Which variables have missing data?

Variables with value of NA:
NonDreaming,
Dreaming,
TotalSleep,
LifeSpan,
Gestation

3. Calculate the mean and standard deviation for TotalSleep.
)
```{r mean and sd}
mean(sleeptable$TotalSleep,na.rm = TRUE)
sd(sleeptable$TotalSleep,na.rm = TRUE)

```

4. Draw a histogram for the BodyWt variable. Note that this variable is highly skewed. Re-draw the histogram on the log scale. Hint: log(sl$BodyWt,base=10). For exta credit, relabel the axes with values at 0.001, 0.01, etc.

```{r histogram of BodyWt log}
hist(log(sleeptable$BodyWt,base=10),xlab = "Body Weight")
```


5. Calculate the ratio of BrainWt to BodyWt and express it as a percentage. Be sure to convert grams to kilograms (or vice versa) before computing the ratio. What animal has the smallest/largest ratio?

```{r calculating ratio and add to data frame}
sleeptable["BrainWt_BodyWt_Ratio"] <- (sleeptable$BrainWt/((sleeptable$BodyWt)*1000))*100
summary(sleeptable$BrainWt_BodyWt_Ratio)
```
```{r smallest and largest}
sleeptable[which.min(sleeptable$BrainWt_BodyWt_Ratio), ]
sleeptable[which.max(sleeptable$BrainWt_BodyWt_Ratio), ]
```
Smallest Ratio = Africanelephant 
Largest Ratio = Groundsquirrel

6. Do bigger animals sleep longer or live longer? Show your results using a correlation coefficient.

```{r correlation}
round(100*cor(sleeptable[,-1],use = "pairwise.complete.obs"),2)
```
  
                  TotalSleep LifeSpan
       BodyWt     -30.72    30.25
Bigger animals live longer. Positive correlation between lifespan and BodyWt

7. Do animals who have high degrees of Predation, Exposure, or Danger sleep less? Show your results using a plot and a trendline.

```{r Total Sleep vs Predation}
plot(sleeptable$Predation,sleeptable$TotalSleep,xlab = "Predation",ylab="Total Sleep")
abline(lm(sleeptable$TotalSleep~sleeptable$Predation))
```
```{r Total Sleep vs Exposure}
plot(sleeptable$Exposure,sleeptable$TotalSleep,xlab = "Exposure",ylab="Total Sleep")
abline(lm(sleeptable$TotalSleep~sleeptable$Exposure))
```
```{r Total Sleep vs Danger}
plot(sleeptable$Danger,sleeptable$TotalSleep,xlab = "Danger",ylab="Total Sleep")
abline(lm(sleeptable$TotalSleep~sleeptable$Danger))
```

Yes, based on the trend line animals with a higher degree of Predation, Exposure, and Danger sleep less

8. Place all the key results into a Word document or PowerPoint presentation, or as a PDF file or as an HTML file. Turn it in by email. Use the subject line: "Introduction to R, (your name), Part 1 Homework"